---
title: "Endowment Percent Change"
author: "Rose Evard"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css', 'template.css')
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "output_html")})
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(kableExtra)
library(plotly)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r}

# make kable table with consistent formatting
make_table <- function(..., title = "", col_names = c("")) {
  title <- paste0("<center><span style = 'font-size:150%;color:black'><b>",
                  title,
                  "</span></b><center>")
   as_tibble(...) %>%
    kbl(caption = title,
        col.names = col_names) %>%
    kable_material() %>%
    row_spec(row=0, background = "#43494C" , color = "white", bold = TRUE)
}

```

```{r}
## Loading in data
endowment_data <- read_rds(here("data", "endowments_by_most_recent_filings.RDS"))
names <- read_csv(here("data", "companies.csv")) %>% 
  ## Dealing with EINs reading in without zero in front
  mutate(EIN = as.character(ein),
         EIN = case_when(
           nchar(EIN) == 8 ~ paste("0", EIN, sep =""),
           TRUE ~ EIN
  )) %>% 
  select(-ein)
```

# Calculating Percent Change

## Including Investments

**(End Year Balance - Beginning Year Balance) / Beginning Year Balance * 100**

> If EYB is larger, positive result. Meaning there was a INCREASE in total funds.  
> If BYB is larger, negative result. Meaning a DECREASE in total funds.  
> If result is above 100, the fund was at least DOUBLED.  
 
```{r, eval = TRUE}
#### USE THIS
## (EYB - BYB)/BYB * 100
## Rose has notes on her choice for this calculation
pct_change_ds <- endowment_data %>%
  filter(!is.na(BeginningYearBalanceAmt)) %>%
  mutate(change = EndYearBalanceAmt - BeginningYearBalanceAmt,
         pct_change = change/BeginningYearBalanceAmt * 100) %>%
  arrange(desc(pct_change)) %>%
  left_join(names, by = "EIN")

```

## REMOVING Investments  

**(End Year Balance - Investment Earnings or Losses Amount - Beginning Year Balance) / Beginning Year Balance * 100**

> Positive result: INCREASE in total funds.  
> Negative result: DECREASE in total funds.  
> If result is above 100, the fund was at least DOUBLED.  

```{r, eval = TRUE}
#### USE THIS
## (EYB - INV - BYB)/BYB * 100
## Rose has notes on her choice for this calculation
pct_change_inv_ds <- endowment_data %>%
  filter(!is.na(BeginningYearBalanceAmt)) %>%
  mutate(InvestmentEarningsOrLossesAmt = ifelse(is.na(InvestmentEarningsOrLossesAmt), 0, InvestmentEarningsOrLossesAmt)) %>%
  mutate(change = EndYearBalanceAmt - InvestmentEarningsOrLossesAmt - BeginningYearBalanceAmt,
         pct_change = change/BeginningYearBalanceAmt * 100) %>%
  arrange(desc(pct_change)) %>%
  left_join(names, by = "EIN")
```


# Summarizing Information  

## With investments  
```{r}
# Basic summary stats
pct_change_ds %>%   
  filter(!is.na(pct_change) & pct_change != Inf) %>%
  summarize(avg_pct_change = mean(pct_change),
            median_pct_change = median(pct_change),
            sd_pct_change = sd(pct_change))

pct_change_ds %>%   
  filter(!is.na(pct_change) & pct_change != Inf) %>%
  group_by(EIN) %>%
  summarize(avg_pct_change = mean(pct_change),
            median_pct_change = median(pct_change),
            sd_pct_change = sd(pct_change))

# Basic histogram summarizing it
ggplot(pct_change_ds, aes(x = pct_change)) +
  geom_histogram(binwidth = 20) + 
  xlab("% Change\n(EYB - BYB) / BYB * 100") + 
  ggtitle(label = "Percentage of Change in Endowment Balance", subtitle = "Red Line indicates 100%") +
  theme_classic() + 
  geom_vline(xintercept = 100, color = "maroon", linetype = "dotted")

## Histogram with cutoff of 200%
ggplot(pct_change_ds, aes(x = pct_change)) +
  geom_histogram(binwidth = 5) + 
  xlab("% Change\n(EYB - BYB) / BYB * 100") + 
  xlim(-100, 200) + 
  ggtitle(label = "Percentage of Change in Endowment Balance") +
  theme_classic()

pct_change_ds %>%
  filter(pct_change != Inf) %>%
  select(organization_name, fiscal_year, pct_change) %>%
  make_table(title = "Percentage of Change in Endowment Balance", col_names = c("Name", "Fiscal Year", "% Change")) %>%
  scroll_box(height = "450px")

pct_change_ds %>%
  filter(pct_change != Inf) %>%
  select(organization_name, fiscal_year, pct_change) %>%
  group_by(fiscal_year) %>%
  summarize(total = n(),
            avg = mean(pct_change),
            med = median(pct_change),
            sd = sd(pct_change)) %>%
  make_table(title = "Percentage of Change in Endowment Balance By Year", col_names = c("Fiscal Year", "Total Companies", "Average % Change", "Median % Change", "Standard Deviation")) %>%
  scroll_box(height = "450px")
```

## REMOVING investments   
```{r}
# Basic summary stats
pct_change_inv_ds %>%   
  filter(!is.na(pct_change) & pct_change != Inf) %>%
  summarize(avg_pct_change = mean(pct_change),
            median_pct_change = median(pct_change),
            sd_pct_change = sd(pct_change))

pct_change_inv_ds %>%   
  filter(!is.na(pct_change) & pct_change != Inf) %>%
  group_by(EIN) %>%
  summarize(avg_pct_change = mean(pct_change),
            median_pct_change = median(pct_change),
            sd_pct_change = sd(pct_change))

# Basic histogram summarizing it
ggplot(pct_change_inv_ds, aes(x = pct_change)) +
  geom_histogram(binwidth = 20) + 
  xlab("% Change\n(EYB - INV - BYB) / BYB * 100") + 
  ggtitle(label = "Percentage of Change in Endowment Balance", subtitle = "Red Line indicates 100%") +
  theme_classic() + 
  geom_vline(xintercept = 100, color = "maroon", linetype = "dotted")

## Histogram with cutoff of 200%
ggplot(pct_change_inv_ds, aes(x = pct_change)) +
  geom_histogram(binwidth = 5) + 
  xlab("% Change\n(EYB - BYB) / BYB * 100") + 
  xlim(-100, 200) + 
  ggtitle(label = "Percentage of Change in Endowment Balance") +
  theme_classic()

pct_change_inv_ds %>%
  filter(pct_change != Inf) %>%
  select(organization_name, fiscal_year, pct_change) %>%
  make_table(title = "Percentage of Change in Endowment Balance", col_names = c("Name", "Fiscal Year", "% Change")) %>%
  scroll_box(height = "450px")

pct_change_inv_ds %>%
  filter(pct_change != Inf) %>%
  select(organization_name, fiscal_year, pct_change) %>%
  group_by(fiscal_year) %>%
  summarize(total = n(),
            avg = mean(pct_change),
            med = median(pct_change),
            sd = sd(pct_change)) %>%
  make_table(title = "Percentage of Change in Endowment Balance By Year", col_names = c("Fiscal Year", "Total Companies", "Average % Change", "Median % Change", "Standard Deviation")) %>%
  scroll_box(height = "450px")
```

# Range of Endowment Percent Change  

## WITH investments  
```{r}
## Ranges of different percent change  
# Reordering by standard deviation of pct_spend down
pct_change_ds_box <- pct_change_ds %>%
  group_by(organization_name) %>%
  filter(pct_change != Inf) %>%
  summarize(sd = sd(pct_change, na.rm = TRUE)) %>%
  right_join(pct_change_ds, by = "organization_name") %>%
  select(organization_name, EIN, pct_change, sd) %>%
  mutate(organization_name = reorder(organization_name, -sd, na.rm = TRUE))

## Unlimited 
box_plot <- ggplot(pct_change_ds_box, aes(x = organization_name, y = pct_change)) + 
  geom_boxplot(aes(color = organization_name), show.legend = FALSE) + 
  geom_point(size = 1, alpha = 0.5) + 
  theme_bw() + 
  labs(title = "Range of Endowment Percent Change per Company",
       x = "Dance Company",
       y = "Percentage of Change in Endowment Balance") + 
  theme(axis.text.x = element_blank()) + 
  geom_hline(yintercept = 100, linetype = "dotted", color = "maroon")
ggplotly(box_plot) %>%
  layout(showlegend = FALSE)

##Limited to 100 for visibility 
box_plot_lim <- ggplot(pct_change_ds_box, aes(x = organization_name, y = pct_change)) + 
  geom_boxplot(aes(color = organization_name), show.legend = FALSE) + 
  geom_point(size = 1, alpha = 0.5) + 
  theme_bw() + 
  labs(title = "Range of Endowment % Change (Max of 100) per Company",
       x = "Dance Company",
       y = "Percentage of Change in Endowment Balance") + 
  theme(axis.text.x = element_blank()) + 
  scale_y_continuous(breaks = scales::breaks_pretty(n = 20),
                     limit = c(-100,100))
ggplotly(box_plot_lim)  %>%
  layout(showlegend = FALSE)


```

## WITHOUT investments  
```{r}
## Ranges of different percent change  
# reorder(organization_name, pull(summarize(spend_down, sd = sd(group_by(spend_down, pct_change)))), na.rm = TRUE)
# Reordering by standard deviation of pct_spend down
pct_change_inv_box <- pct_change_inv_ds %>%
  group_by(organization_name) %>%
  filter(pct_change != Inf) %>%
  summarize(sd = sd(pct_change, na.rm = TRUE)) %>%
  right_join(pct_change_ds, by = "organization_name") %>%
  select(organization_name, EIN, pct_change, sd) %>%
  mutate(organization_name = reorder(organization_name, -sd, na.rm = TRUE))

## Unlimited 
box_plot_inv <- ggplot(pct_change_inv_box, aes(x = organization_name, y = pct_change)) + 
  geom_boxplot(aes(color = organization_name), show.legend = FALSE) + 
  geom_point(size = 1, alpha = 0.5) + 
  theme_bw() + 
  labs(title = "Range of Endowment Percent Change per Company",
       x = "Dance Company",
       y = "Percentage of Change in Endowment Balance") + 
  theme(axis.text.x = element_blank()) + 
  geom_hline(yintercept = 100, linetype = "dotted", color = "maroon")
ggplotly(box_plot_inv) %>%
  layout(showlegend = FALSE)

##Limited to 100 for visibility 
box_plot_inv_lim <- ggplot(pct_change_inv_box, aes(x = organization_name, y = pct_change)) + 
  geom_boxplot(aes(color = organization_name), show.legend = FALSE) + 
  geom_point(size = 1, alpha = 0.5) + 
  theme_bw() + 
  labs(title = "Range of Endowment % Change (Max of 100) per Company",
       x = "Dance Company",
       y = "Percentage of Change in Endowment Balance") + 
  theme(axis.text.x = element_blank()) + 
  scale_y_continuous(breaks = scales::breaks_pretty(n = 20),
                     limit = c(-100,100))
ggplotly(box_plot_inv_lim)  %>%
  layout(showlegend = FALSE)


```


# Percent Change over Time   
## WITH investments  

```{r}
## Retrieving median values  
pct_change_med <- pct_change_ds %>%
  group_by(fiscal_year) %>%
  summarize(median = median(pct_change, na.rm = TRUE)) %>%
  mutate(organization_name = "Median")
```

```{r}  
## Spend Down over Time
pct_change_plot <- pct_change_ds %>%
  ggplot(aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = organization_name)) +
  geom_line(alpha = 0.5) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance",
       subtitle = "By Fiscal Year") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20)) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "gray") +
  geom_point(data = pct_change_med, aes(x = fiscal_year, y = median), color = "black", size = 1)
ggplotly(pct_change_plot) 
```

```{r}
##Plot with Y scale between -100 and 100 
limited_scale <- pct_change_ds %>%
  ggplot(aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = organization_name)) +
  geom_line(show.legend = FALSE, alpha = 0.5) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance (max 100)",
       subtitle = "By Fiscal Year") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20),
                     limits = c(-100, 100))  +
  geom_point(data = pct_change_med, aes(x = fiscal_year, y = median), color = "black", size = 1)
ggplotly(limited_scale)
```

## WITHOUT investments   

```{r}
## Retrieving median values  
pct_change_inv_med <- pct_change_inv_ds %>%
  group_by(fiscal_year) %>%
  summarize(median = median(pct_change, na.rm = TRUE)) %>%
  mutate(organization_name = "Median")
```


```{r}  
## Spend Down over Time
pct_change_inv_plot <- pct_change_inv_ds %>%
  ggplot(aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = organization_name)) +
  geom_line(alpha = 0.5) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance",
       subtitle = "By Fiscal Year") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20)) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "gray")  +
  geom_point(data = pct_change_inv_med, aes(x = fiscal_year, y = median), color = "black", size = 1)
ggplotly(pct_change_inv_plot) 
```

```{r}
##Plot with Y scale between -100 and 100 
limited_scale_inv <- pct_change_inv_ds %>%
  ggplot(aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = organization_name)) +
  geom_line(show.legend = FALSE, alpha = 0.5) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance (max 100)",
       subtitle = "By Fiscal Year") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20),
                     limits = c(-100, 100))  +
  geom_point(data = pct_change_inv_med, aes(x = fiscal_year, y = median), color = "black", size = 0.5)
ggplotly(limited_scale_inv)
```

# Comparing within-year percent change to S&P 500  

Adding S&P index  

Altering plots to be plotting date-wise so the X-axes can be the same 

```{r}
## Adding end date 
source(here("GET_VARS.R"))
##Specifically reading in Tax Period End Date
files <- dir( here("ballet_990_released_20230208"),
              full.names = TRUE)
taxperiod_data <- map_df(files, ~
                    get_df(variables =  c("//Return//ReturnHeader//TaxPeriodEndDt"),
                           filename = .x
                           )) 

taxperiod_data$fiscal_year = as.numeric(as.character(taxperiod_data$fiscal_year))

pct_change_rtd <- pct_change_ds %>%
  left_join(taxperiod_data, by = c("EIN", "fiscal_year"))

pct_change_inv_rtd <- pct_change_inv_ds %>%
  left_join(taxperiod_data, by = c("EIN", "fiscal_year"))

## Imputing return_date from N/A 
# Assuming that month/day is the same, fiscal year is one year prior 
pct_change_rtd <- pct_change_rtd %>%
  group_by(EIN) %>%
  arrange(desc(EIN), desc(fiscal_year)) %>%
  mutate(TaxPeriodEndDt= ifelse(is.na(TaxPeriodEndDt), 
                    paste(fiscal_year, substring(lag(TaxPeriodEndDt), 5, 10), sep = ""), 
                    as.character(TaxPeriodEndDt))) 

pct_change_inv_rtd <- pct_change_inv_rtd %>%
  group_by(EIN) %>%
  arrange(desc(EIN), desc(fiscal_year)) %>%
  mutate(TaxPeriodEndDt= ifelse(is.na(TaxPeriodEndDt), 
                    paste(fiscal_year, substring(lag(TaxPeriodEndDt), 5, 10), sep = ""), 
                    as.character(TaxPeriodEndDt))) 
#Produces some NAs due to vectorization
  
# Rids all '2010NA' etcs.  
# Only have 11 years, so at max will run 11 times 
for (i in 1:11) {
  pct_change_rtd <- pct_change_rtd %>%
    mutate(TaxPeriodEndDt= ifelse(grepl("NA", TaxPeriodEndDt), 
                    paste(fiscal_year, substring(lag(TaxPeriodEndDt), 5, 10), sep = ""), 
                    as.character(TaxPeriodEndDt)))
  
    pct_change_inv_rtd <- pct_change_inv_rtd %>%
    mutate(TaxPeriodEndDt= ifelse(grepl("NA", TaxPeriodEndDt), 
                    paste(fiscal_year, substring(lag(TaxPeriodEndDt), 5, 10), sep = ""), 
                    as.character(TaxPeriodEndDt)))
}

pct_change_rtd$TaxPeriodEndDt = as.Date(pct_change_rtd$TaxPeriodEndDt, format = "%Y-%m-%d")
pct_change_inv_rtd$TaxPeriodEndDt = as.Date(pct_change_inv_rtd$TaxPeriodEndDt, format = "%Y-%m-%d")

pct_change_rtd <- ungroup(pct_change_rtd)
pct_change_inv_rtd <- ungroup(pct_change_inv_rtd)
```

```{r}  
## Gets just values from beginnings of the months 
sp_begin <- read_csv(here('data','SP500.csv')) %>%
  rename_with(tolower) %>%
  mutate(sp500 = as.numeric(sp500)) %>%
  filter(!is.na(sp500)) %>%
  mutate(month = month(date),
         year = year(date)) %>%
  group_by(month,year) %>%
  arrange(month) %>%
 slice_min(n=1,order_by = date) %>%
  ungroup() %>%
  mutate(type = "START")

##Values from the end of the month
sp_end <- read_csv(here('data','SP500.csv')) %>%
  rename_with(tolower) %>%
  mutate(sp500 = as.numeric(sp500)) %>%
  filter(!is.na(sp500)) %>%
  mutate(month = month(date),
         year = year(date)) %>%
  group_by(month,year) %>%
  arrange(month) %>%
 slice_max(n=1,order_by = date) %>%
  ungroup() %>%
  mutate(type = "END")

sp_pct_chg <- sp_begin %>%
  bind_rows(sp_end) %>%
  pivot_wider(names_from = type, values_from = c(date, sp500)) %>%
  mutate(pct_change = (sp500_END - sp500_START)/sp500_START * 100)
```

## WITH investments  
```{r}
## SPend down over time with new X axis 
## Median by year 
pct_change_med <- pct_change_rtd %>%
  mutate(fiscal_year = paste(fiscal_year, "-01-01", sep = ""),
         fiscal_year = as.Date(fiscal_year, "%Y-%m-%d")) %>%
  group_by(fiscal_year) %>%
  summarize(median = median(pct_change, na.rm = TRUE)) %>%
  mutate(organization_name = "Median")

## Spend Down over Time
pct_change_rtd %>%
  filter(pct_change != Inf) %>%
  ggplot(aes(x = TaxPeriodEndDt, y = pct_change)) +
  geom_line(aes(group = organization_name, color = organization_name), alpha = 0.5, show.legend = FALSE) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance",
       subtitle = "Black Dot: Median. Black line: S&P500 % change in a month") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 10),
                     limits = c(-100, 100)) +
  scale_x_date(breaks = scales::pretty_breaks(n = 20)) +
  geom_point(data = pct_change_med, aes(x = fiscal_year, y = median), color = "black", size = 1) + 
  geom_line(data = sp_pct_chg, aes(x = date_END, y = pct_change), color = "black", show.legend = FALSE)
```

Issue with this table: PCT change in S&P 500 is calculated month-by-month, while PCT change in endowment is calculated across a fiscal year.  

```{r}
## GETTING PERCENT CHANGE OF FISCAL YEARS THAT EXIST IN THE DATASET
## All 'end' dates of fiscal years
date_ends <- pct_change_rtd %>%
  pull(TaxPeriodEndDt) %>%
  unique()

sp_end_month <- read_csv(here('data','SP500.csv')) %>%
  rename_with(tolower) %>%
  mutate(sp500 = as.numeric(sp500)) %>%
  filter(!is.na(sp500)) %>%
  filter(date %in% date_ends) %>%
  select(date_end = date, sp500_end = sp500) %>%
  mutate(month = month(date_end),
         year = year(date_end),
         year_last = year(date_end) - 1)

sp_start_month <- read_csv(here('data','SP500.csv')) %>%
  rename_with(tolower) %>%
  mutate(sp500 = as.numeric(sp500)) %>%
  filter(!is.na(sp500)) %>%
  select(date_start = date, sp500_start = sp500) %>%
  mutate(month = month(date_start),
         year = year(date_start)) %>%
  group_by(month,year) %>%
  arrange(month) %>%
 slice_min(n=1,order_by = date_start) %>%
  ungroup()

sp_change_years <- sp_end_month %>%
  left_join(sp_start_month, by = c("month", "year_last" = "year")) %>%
  mutate(pct_change = (sp500_end - sp500_start)/sp500_start * 100)
```

```{r}
pct_change_rtd %>%
  mutate(month = as.numeric(substring(TaxPeriodEndDt, 6, 7))) %>%
  filter(pct_change != Inf) %>%
  left_join(sp_change_years, by = c("TaxPeriodEndDt" = "date_end")) %>%
  select(organization_name, TaxPeriodEndDt, pct_change.x, pct_change.y) %>%
  arrange(desc(abs(pct_change.x))) %>%
  make_table(title = "Percentage Change of Endowment and S&P 500", 
             col_names = c("Organization Name", "Fiscal Year End Date", "Endowment Percent Change", "S&P 500 Percent Change")) %>%
  scroll_box(height = "450px")
```


## WITHOUT investments 
```{r}
## SPend down over time with new X axis 
## Median by year 
pct_change_inv_med <- pct_change_inv_rtd %>%
  mutate(fiscal_year = paste(fiscal_year, "-01-01", sep = ""),
         fiscal_year = as.Date(fiscal_year, "%Y-%m-%d")) %>%
  group_by(fiscal_year) %>%
  summarize(median = median(pct_change, na.rm = TRUE)) %>%
  mutate(organization_name = "Median")

## Spend Down over Time
pct_change_inv_rtd %>%
  filter(pct_change != Inf) %>%
  ggplot(aes(x = TaxPeriodEndDt, y = pct_change)) +
  geom_line(aes(group = organization_name, color = organization_name), alpha = 0.5, show.legend = FALSE) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance",
       subtitle = "Black Dot: Median. Black line: S&P500 % change in a month") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 10),
                     limits = c(-100, 100)) +
  scale_x_date(breaks = scales::pretty_breaks(n = 20)) +
  geom_point(data = pct_change_inv_med, aes(x = fiscal_year, y = median), color = "black", size = 1) + 
  geom_line(data = sp_pct_chg, aes(x = date_END, y = pct_change), color = "black", show.legend = FALSE)
```


> NTS: Add table 

# Breaking Companies up by Behavior   
Using only the dataset normalized for investments, as that has more consistency.  

```{r}
# Adding var for "Consistency"  

## Between mean +/- SD
#inv_consist <- pct_change_inv_ds %>%
 # group_by(organization_name) %>%
  #filter(pct_change != Inf & pct_change != -Inf) %>%
#  summarize(mean = mean(pct_change), 
 #           sd = sd(pct_change)) %>%
 # mutate(consistent = ifelse(between(mean + sd, -7, 7) | between(mean - sd, -7, 7), TRUE, FALSE))

## Checking if any value is outside of -7:7, can do with min/max
inv_consist <- pct_change_inv_ds %>%
  group_by(organization_name) %>%
  filter(pct_change != Inf & pct_change != -Inf) %>%
  summarize(min = min(pct_change), 
            max = max(pct_change),
            sd = sd(pct_change)) %>%
  mutate(consistent = ifelse(between(min + sd, -7, 7) & between(max - sd, -7, 7), TRUE, FALSE),
         #Manual inspection showed BQC should be marked as consistent but just didnt quite make it math-wise
         consistent = ifelse(organization_name == "Ballet Quad Cities", TRUE, consistent))

## Adding consistency 
inv_consist_ds <- inv_consist %>% 
  select(organization_name, consistent) %>% 
  right_join(pct_change_inv_ds, by = "organization_name") %>%
  filter(pct_change != Inf & pct_change != -Inf)

```

```{r}
##plottinng
test <- ggplot(inv_consist_ds, aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = consistent)) +
  geom_line(show.legend = TRUE, alpha = 0.3) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance (max 100)",
       subtitle = "By Fiscal Year") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20),
                     limits = c(-100, 100))  +
  geom_point(data = pct_change_inv_med, aes(x = fiscal_year, y = median), color = "black", size = 0.5) + 
  geom_hline(yintercept = 7, alpha = .5, linetype = "dotted") + 
  geom_hline(yintercept = -7, alpha = .5, linetype = "dotted") 

test
ggplotly(test)

inv_box <- inv_consist_ds %>%
  group_by(organization_name) %>%
  filter(pct_change != Inf) %>%
  summarize(sd = sd(pct_change, na.rm = TRUE)) %>%
  right_join(inv_consist_ds, by = "organization_name") %>%
  select(organization_name, EIN, pct_change, sd, consistent) %>%
  mutate(organization_name = reorder(organization_name, -sd, na.rm = TRUE))

## Unlimited 
box_plot <- ggplot(inv_box, aes(x = organization_name, y = pct_change)) + 
  geom_boxplot(aes(color = consistent), show.legend = FALSE) + 
  geom_point(size = 1, alpha = 0.5) + 
  theme_bw() + 
  labs(title = "Range of Endowment Percent Change per Company",
       x = "Dance Company",
       y = "Percentage of Change in Endowment Balance") + 
  theme(axis.text.x = element_blank()) + 
  geom_hline(yintercept = 100, linetype = "dotted", color = "maroon")
ggplotly(box_plot) %>%
  layout(showlegend = FALSE)

##Limited to 100 for visibility 
ggplot(inv_box, aes(x = organization_name, y = pct_change)) + 
  geom_boxplot(aes(color = consistent), show.legend = FALSE) + 
  geom_point(size = 1, alpha = 0.2) + 
  theme_bw() + 
  labs(title = "Range of Endowment % Change (Max of 100) per Company",
       x = "Dance Company",
       y = "Percentage of Change in Endowment Balance") + 
  theme(axis.text.x = element_blank()) + 
  scale_y_continuous(breaks = scales::breaks_pretty(n = 20),
                     limit = c(-100,100))
```

## Examining the Two Classes

### Lists of Companies within Each Class
```{r}
## Names in a table  
# CONSISTENT ones  
inv_consist %>% 
  filter(consistent == TRUE) %>%
  select(-consistent) %>%
  make_table(title = "Companies with Consistent Endowment Changes", col_names = c("Organization Name", "Minimum Change", "Maximum Change", "Standard Deviation")) %>%
  scroll_box(height = "450px")
```
 
There are `r inv_consist %>% filter(consistent == TRUE) %>% nrow()` companies which have 'consistent' endowment percent change. 


```{r}
## Names in a table  
# INCONSISTENT ones  
inv_consist %>% 
  filter(consistent == FALSE) %>%
  select(-consistent) %>%
  make_table(title = "Companies with Inconsistent Endowment Changes", col_names = c("Organization Name", "Minimum Change", "Maximum Change", "Standard Deviation")) %>%
  scroll_box(height = "450px")
```

There are `r inv_consist %>% filter(consistent == FALSE) %>% nrow()` companies which have 'inconsistent' endowment percent change. 

### Size of Companies in Each Class   

```{r}
##Getting employee data
source(here("GET_VARS.R"))
##Specifically reading in employee data
files <- dir( here("ballet_990_released_20230208"),
              full.names = TRUE)
employ_data <- map_df(files, ~
                    get_df(variables =  c("//Return//ReturnData//TotalEmployeeCnt"),
                           filename = .x
                           )) %>%
    mutate(TotalEmployeeCnt = as.numeric(TotalEmployeeCnt)) %>%
  left_join(names, by = "EIN")
```

```{r}
## Most recent years for each company
employ_data$fiscal_year = as.numeric(as.character(employ_data$fiscal_year))
most_recent_yrs <- employ_data %>%
  group_by(organization_name) %>%
  summarize(recent_year = max(fiscal_year, na.rm = TRUE)) %>%
  left_join(employ_data, by = c("organization_name", "recent_year" = "fiscal_year"))

## Limiting to ONLY most recent year 
most_recent_ds <- inv_consist_ds %>%
  left_join(most_recent_yrs, by = "organization_name") %>%
  filter(fiscal_year == recent_year)
```

```{r}
## Consistent  
most_recent_ds %>%
  filter(consistent == TRUE) %>%
  select(organization_name, TotalEmployeeCnt, EndYearBalanceAmt, recent_year) %>%
  make_table("Size of Consistent Companies, by Employees and Endowment", col_names = c("Organization Name", "Total Employee Count", "End Year Balance Amount", "Fiscal Year")) %>%
  scroll_box(height = "450px")
```


```{r}
## Inconsistent  
most_recent_ds %>%
  filter(consistent == FALSE) %>%
  filter(!is.na(TotalEmployeeCnt)) %>%
  select(organization_name, TotalEmployeeCnt, EndYearBalanceAmt, recent_year) %>%
  make_table("Size of Inconsistent Companies, by Employees and Endowment", col_names = c("Organization Name", "Total Employee Count", "End Year Balance Amount", "Fiscal Year")) %>%
  scroll_box(height = "450px")
```

```{r}

## Labels for facet 
consistent_labels <- c(
  `TRUE` = "Consistent",
  `FALSE` = "Inconsistent"
)

##Histograms
most_recent_ds %>%
  ggplot(aes(x = TotalEmployeeCnt)) +
  geom_histogram(binwidth = 60, color = "black", fill = "gray") +
  facet_wrap(~consistent, labeller = as_labeller(consistent_labels)) + 
  theme_bw() + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 90),
        strip.text = element_text(face="bold",size = 10),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
    scale_x_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 10)) +
  labs( title = "Size of Companies by Employee Total")

most_recent_ds %>%
  ggplot(aes(x = EndYearBalanceAmt)) +
  geom_histogram(color = "black", fill = "gray") + 
  facet_wrap(~consistent, labeller = as_labeller(consistent_labels)) + 
  theme_bw() + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 90),
        strip.text = element_text(face="bold",size = 10),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
    scale_x_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 10)) +
  labs( title = "Size of Companies by End Year Balance Amount")

most_recent_ds %>%
  filter(EndYearBalanceAmt < 20000000) %>%
  ggplot(aes(x = EndYearBalanceAmt)) +
  geom_histogram(color = "black", fill = "gray") + 
  facet_wrap(~consistent, labeller = as_labeller(consistent_labels)) + 
  theme_bw() +
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 90),
        strip.text = element_text(face="bold",size = 10),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
    scale_x_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 10)) +
  labs( title = "Size of Companies by End Year Balance Amount, Limit $20,000,000")
```

# Total Revenue  

```{r}
##Getting Total Revenue data
source(here("GET_VARS.R"))
##Specifically reading in employee data
files <- dir( here("ballet_990_released_20230208"),
              full.names = TRUE)
revenue_data <- map_df(files, ~
                    get_df(variables =  c("//Return//ReturnData//CYTotalRevenueAmt"),
                           filename = .x
                           )) %>%
    mutate(CYTotalRevenueAmt = as.numeric(CYTotalRevenueAmt)) %>%
  left_join(names, by = "EIN")

revenue_data$fiscal_year = as.numeric(as.character(revenue_data$fiscal_year))
```

## Total Revenue vs Pct Change- any pattern?  

```{r}
ggplot(revenue_data, aes(x = CYTotalRevenueAmt)) + 
  geom_histogram()
```

```{r}
pct_change_inv_ds %>%
  select(-dance_styles) %>%
  left_join(revenue_data, by = c("EIN", "fiscal_year", "organization_name")) %>%
  ggplot(aes(x = CYTotalRevenueAmt, y = pct_change)) + 
  geom_point() + 
  theme_bw() + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20),
                     limits = c(-200, 800))
```


# Percent Change Within the Pandemic   
## WITH investments 
```{r}  
## Pandemic Years
pct_change_plot <- pct_change_ds %>%
  filter(fiscal_year %in% c("2018", "2019", "2020", "2021", "2022")) %>% 
  ggplot(aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = organization_name)) +
  geom_line(show.legend = FALSE, alpha = 0.5) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance",
       subtitle = "Within Pandemic Years") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20)) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "gray")
ggplotly(pct_change_plot) 
```

```{r}
## Table of available in-pandemic data 
pct_change_ds %>%
  filter(fiscal_year %in% c("2019", "2020", "2021", "2022")) %>%
  select(organization_name, pct_change, fiscal_year) %>%
  arrange(desc(fiscal_year)) %>%
  make_table(title = "Percentage of Change in Endowment Balance within Pandemic Years", col_names = c("Name", "% Change", "Year")) %>%
  scroll_box(height = "450px")
```

```{r}
pct_change_ds %>%
  filter(fiscal_year %in% c("2019", "2020", "2021", "2022")) %>%
  select(organization_name, pct_change, fiscal_year) %>%
  group_by(fiscal_year) %>%
  summarize(total_in_year = n())
```

## WITHOUT investments 
```{r}  
## Pandemic Years
pct_change_plot_inv <- pct_change_inv_ds %>%
  filter(fiscal_year %in% c("2018", "2019", "2020", "2021", "2022")) %>% 
  ggplot(aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = organization_name)) +
  geom_line(show.legend = FALSE, alpha = 0.5) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance",
       subtitle = "Within Pandemic Years") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20)) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "gray")
ggplotly(pct_change_plot_inv) 
```

```{r}
## Table of available in-pandemic data 
pct_change_inv_ds %>%
  filter(fiscal_year %in% c("2019", "2020", "2021", "2022")) %>%
  select(organization_name, pct_change, fiscal_year) %>%
  arrange(desc(fiscal_year)) %>%
  make_table(title = "Percentage of Change in Endowment Balance within Pandemic Years", col_names = c("Name", "% Change", "Year")) %>%
  scroll_box(height = "450px")
```

```{r}
pct_change_inv_ds %>%
  filter(fiscal_year %in% c("2019", "2020", "2021", "2022")) %>%
  select(organization_name, pct_change, fiscal_year) %>%
  group_by(fiscal_year) %>%
  summarize(total_in_year = n())
```

