---
title: "Endowment Percent Change"
author: "Rose Evard"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css', 'template.css')
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "output_html")})
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(kableExtra)
library(plotly)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r}

# make kable table with consistent formatting
make_table <- function(..., title = "", col_names = c("")) {
  title <- paste0("<center><span style = 'font-size:150%;color:black'><b>",
                  title,
                  "</span></b><center>")
   as_tibble(...) %>%
    kbl(caption = title,
        col.names = col_names) %>%
    kable_material() %>%
    row_spec(row=0, background = "#43494C" , color = "white", bold = TRUE)
}

```

```{r}
## Loading in data
endowment_data <- read_rds(here("data", "endowments_by_most_recent_filings.RDS"))
names <- read_csv(here("data", "companies.csv")) %>% 
  ## Dealing with EINs reading in without zero in front
  mutate(EIN = as.character(ein),
         EIN = case_when(
           nchar(EIN) == 8 ~ paste("0", EIN, sep =""),
           TRUE ~ EIN
  )) %>% 
  select(-ein)
```

# Calculating Percent Change

## Including Investments

**(End Year Balance - Beginning Year Balance) / Beginning Year Balance * 100**

> If EYB is larger, positive result. Meaning there was a INCREASE in total funds.  
> If BYB is larger, negative result. Meaning a DECREASE in total funds.  
> If result is above 100, the fund was at least DOUBLED.  
 
```{r, eval = TRUE}
#### USE THIS
## (EYB - BYB)/BYB * 100
## Rose has notes on her choice for this calculation
pct_change_ds <- endowment_data %>%
  filter(!is.na(BeginningYearBalanceAmt)) %>%
  mutate(change = EndYearBalanceAmt - BeginningYearBalanceAmt,
         pct_change = change/BeginningYearBalanceAmt * 100) %>%
  arrange(desc(pct_change)) %>%
  left_join(names, by = "EIN")

```

## REMOVING Investments  

**(End Year Balance - Investment Earnings or Losses Amount - Beginning Year Balance) / Beginning Year Balance * 100**

> Positive result: INCREASE in total funds.  
> Negative result: DECREASE in total funds.  
> If result is above 100, the fund was at least DOUBLED.  

```{r, eval = TRUE}
#### USE THIS
## (EYB - INV - BYB)/BYB * 100
## Rose has notes on her choice for this calculation
pct_change_inv_ds <- endowment_data %>%
  filter(!is.na(BeginningYearBalanceAmt)) %>%
  mutate(InvestmentEarningsOrLossesAmt = ifelse(is.na(InvestmentEarningsOrLossesAmt), 0, InvestmentEarningsOrLossesAmt)) %>%
  mutate(change = EndYearBalanceAmt - InvestmentEarningsOrLossesAmt - BeginningYearBalanceAmt,
         pct_change = change/BeginningYearBalanceAmt * 100) %>%
  arrange(desc(pct_change)) %>%
  left_join(names, by = "EIN")
```


# Summarizing Information  

## With investments  
```{r}
# Basic summary stats
pct_change_ds %>%   
  filter(!is.na(pct_change) & pct_change != Inf) %>%
  summarize(avg_pct_change = mean(pct_change),
            median_pct_change = median(pct_change),
            sd_pct_change = sd(pct_change))

pct_change_ds %>%   
  filter(!is.na(pct_change) & pct_change != Inf) %>%
  group_by(EIN) %>%
  summarize(avg_pct_change = mean(pct_change),
            median_pct_change = median(pct_change),
            sd_pct_change = sd(pct_change))

# Basic histogram summarizing it
ggplot(pct_change_ds, aes(x = pct_change)) +
  geom_histogram(binwidth = 20) + 
  xlab("% Change\n(EYB - BYB) / BYB * 100") + 
  ggtitle(label = "Percentage of Change in Endowment Balance", subtitle = "Red Line indicates 100%") +
  theme_classic() + 
  geom_vline(xintercept = 100, color = "maroon", linetype = "dotted")

## Histogram with cutoff of 200%
ggplot(pct_change_ds, aes(x = pct_change)) +
  geom_histogram(binwidth = 5) + 
  xlab("% Change\n(EYB - BYB) / BYB * 100") + 
  xlim(-100, 200) + 
  ggtitle(label = "Percentage of Change in Endowment Balance") +
  theme_classic()

pct_change_ds %>%
  filter(pct_change != Inf) %>%
  select(organization_name, fiscal_year, pct_change) %>%
  make_table(title = "Percentage of Change in Endowment Balance", col_names = c("Name", "Fiscal Year", "% Change")) %>%
  scroll_box(height = "450px")

pct_change_ds %>%
  filter(pct_change != Inf) %>%
  select(organization_name, fiscal_year, pct_change) %>%
  group_by(fiscal_year) %>%
  summarize(total = n(),
            avg = mean(pct_change),
            med = median(pct_change),
            sd = sd(pct_change)) %>%
  make_table(title = "Percentage of Change in Endowment Balance By Year", col_names = c("Fiscal Year", "Total Companies", "Average % Change", "Median % Change", "Standard Deviation")) %>%
  scroll_box(height = "450px")
```

## REMOVING investments   
```{r}
# Basic summary stats
pct_change_inv_ds %>%   
  filter(!is.na(pct_change) & pct_change != Inf) %>%
  summarize(avg_pct_change = mean(pct_change),
            median_pct_change = median(pct_change),
            sd_pct_change = sd(pct_change))

pct_change_inv_ds %>%   
  filter(!is.na(pct_change) & pct_change != Inf) %>%
  group_by(EIN) %>%
  summarize(avg_pct_change = mean(pct_change),
            median_pct_change = median(pct_change),
            sd_pct_change = sd(pct_change))

# Basic histogram summarizing it
ggplot(pct_change_inv_ds, aes(x = pct_change)) +
  geom_histogram(binwidth = 20) + 
  xlab("% Change\n(EYB - INV - BYB) / BYB * 100") + 
  ggtitle(label = "Percentage of Change in Endowment Balance", subtitle = "Red Line indicates 100%") +
  theme_classic() + 
  geom_vline(xintercept = 100, color = "maroon", linetype = "dotted")

## Histogram with cutoff of 200%
ggplot(pct_change_inv_ds, aes(x = pct_change)) +
  geom_histogram(binwidth = 5) + 
  xlab("% Change\n(EYB - BYB) / BYB * 100") + 
  xlim(-100, 200) + 
  ggtitle(label = "Percentage of Change in Endowment Balance") +
  theme_classic()

pct_change_inv_ds %>%
  filter(pct_change != Inf) %>%
  select(organization_name, fiscal_year, pct_change) %>%
  make_table(title = "Percentage of Change in Endowment Balance", col_names = c("Name", "Fiscal Year", "% Change")) %>%
  scroll_box(height = "450px")

pct_change_inv_ds %>%
  filter(pct_change != Inf) %>%
  select(organization_name, fiscal_year, pct_change) %>%
  group_by(fiscal_year) %>%
  summarize(total = n(),
            avg = mean(pct_change),
            med = median(pct_change),
            sd = sd(pct_change)) %>%
  make_table(title = "Percentage of Change in Endowment Balance By Year", col_names = c("Fiscal Year", "Total Companies", "Average % Change", "Median % Change", "Standard Deviation")) %>%
  scroll_box(height = "450px")
```

# Range of Endowment Percent Change  

## WITH investments  
```{r}
## Ranges of different percent change  
# Reordering by standard deviation of pct_spend down
pct_change_ds_box <- pct_change_ds %>%
  group_by(organization_name) %>%
  filter(pct_change != Inf) %>%
  summarize(sd = sd(pct_change, na.rm = TRUE)) %>%
  right_join(pct_change_ds, by = "organization_name") %>%
  select(organization_name, EIN, pct_change, sd) %>%
  mutate(organization_name = reorder(organization_name, -sd, na.rm = TRUE))

## Unlimited 
box_plot <- ggplot(pct_change_ds_box, aes(x = organization_name, y = pct_change)) + 
  geom_boxplot(aes(color = organization_name), show.legend = FALSE) + 
  geom_point(size = 1, alpha = 0.5) + 
  theme_bw() + 
  labs(title = "Range of Endowment Percent Change per Company",
       x = "Dance Company",
       y = "Percentage of Change in Endowment Balance") + 
  theme(axis.text.x = element_blank()) + 
  geom_hline(yintercept = 100, linetype = "dotted", color = "maroon")
ggplotly(box_plot) %>%
  layout(showlegend = FALSE)

##Limited to 100 for visibility 
box_plot_lim <- ggplot(pct_change_ds_box, aes(x = organization_name, y = pct_change)) + 
  geom_boxplot(aes(color = organization_name), show.legend = FALSE) + 
  geom_point(size = 1, alpha = 0.5) + 
  theme_bw() + 
  labs(title = "Range of Endowment % Change (Max of 100) per Company",
       x = "Dance Company",
       y = "Percentage of Change in Endowment Balance") + 
  theme(axis.text.x = element_blank()) + 
  scale_y_continuous(breaks = scales::breaks_pretty(n = 20),
                     limit = c(-100,100))
ggplotly(box_plot_lim)  %>%
  layout(showlegend = FALSE)


```

## WITHOUT investments  
```{r}
## Ranges of different percent change  
# reorder(organization_name, pull(summarize(spend_down, sd = sd(group_by(spend_down, pct_change)))), na.rm = TRUE)
# Reordering by standard deviation of pct_spend down
pct_change_inv_box <- pct_change_inv_ds %>%
  group_by(organization_name) %>%
  filter(pct_change != Inf) %>%
  summarize(sd = sd(pct_change, na.rm = TRUE)) %>%
  right_join(pct_change_ds, by = "organization_name") %>%
  select(organization_name, EIN, pct_change, sd) %>%
  mutate(organization_name = reorder(organization_name, -sd, na.rm = TRUE))

## Unlimited 
box_plot_inv <- ggplot(pct_change_inv_box, aes(x = organization_name, y = pct_change)) + 
  geom_boxplot(aes(color = organization_name), show.legend = FALSE) + 
  geom_point(size = 1, alpha = 0.5) + 
  theme_bw() + 
  labs(title = "Range of Endowment Percent Change per Company",
       x = "Dance Company",
       y = "Percentage of Change in Endowment Balance") + 
  theme(axis.text.x = element_blank()) + 
  geom_hline(yintercept = 100, linetype = "dotted", color = "maroon")
ggplotly(box_plot_inv) %>%
  layout(showlegend = FALSE)

##Limited to 100 for visibility 
box_plot_inv_lim <- ggplot(pct_change_inv_box, aes(x = organization_name, y = pct_change)) + 
  geom_boxplot(aes(color = organization_name), show.legend = FALSE) + 
  geom_point(size = 1, alpha = 0.5) + 
  theme_bw() + 
  labs(title = "Range of Endowment % Change (Max of 100) per Company",
       x = "Dance Company",
       y = "Percentage of Change in Endowment Balance") + 
  theme(axis.text.x = element_blank()) + 
  scale_y_continuous(breaks = scales::breaks_pretty(n = 20),
                     limit = c(-100,100))
ggplotly(box_plot_inv_lim)  %>%
  layout(showlegend = FALSE)


```


# Percent Change over Time   
## WITH investments  

```{r}
## Retrieving median values  
pct_change_med <- pct_change_ds %>%
  group_by(fiscal_year) %>%
  summarize(median = median(pct_change, na.rm = TRUE)) %>%
  mutate(organization_name = "Median")
```

```{r}  
## Spend Down over Time
pct_change_plot <- pct_change_ds %>%
  ggplot(aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = organization_name)) +
  geom_line(alpha = 0.5) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance",
       subtitle = "By Fiscal Year") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20)) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "gray") +
  geom_point(data = pct_change_med, aes(x = fiscal_year, y = median), color = "black", size = 1)
ggplotly(pct_change_plot) 
```

```{r}
##Plot with Y scale between -100 and 100 
limited_scale <- pct_change_ds %>%
  ggplot(aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = organization_name)) +
  geom_line(show.legend = FALSE, alpha = 0.5) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance (max 100)",
       subtitle = "By Fiscal Year") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20),
                     limits = c(-100, 100))  +
  geom_point(data = pct_change_med, aes(x = fiscal_year, y = median), color = "black", size = 1)
ggplotly(limited_scale)
```

## WITHOUT investments   

```{r}
## Retrieving median values  
pct_change_inv_med <- pct_change_inv_ds %>%
  group_by(fiscal_year) %>%
  summarize(median = median(pct_change, na.rm = TRUE)) %>%
  mutate(organization_name = "Median")
```


```{r}  
## Spend Down over Time
pct_change_inv_plot <- pct_change_inv_ds %>%
  ggplot(aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = organization_name)) +
  geom_line(alpha = 0.5) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance",
       subtitle = "By Fiscal Year") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20)) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "gray")  +
  geom_point(data = pct_change_inv_med, aes(x = fiscal_year, y = median), color = "black", size = 1)
ggplotly(pct_change_inv_plot) 
```

```{r}
##Plot with Y scale between -100 and 100 
limited_scale_inv <- pct_change_inv_ds %>%
  ggplot(aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = organization_name)) +
  geom_line(show.legend = FALSE, alpha = 0.5) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance (max 100)",
       subtitle = "By Fiscal Year") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20),
                     limits = c(-100, 100))  +
  geom_point(data = pct_change_inv_med, aes(x = fiscal_year, y = median), color = "black", size = 0.5)
ggplotly(limited_scale_inv)
```

# Breaking Companies up by Behavior   
Using only the dataset normalized for investments, as that has more consistency.  

```{r}
# Adding var for "Consistency"  

## Between mean +/- SD
#inv_consist <- pct_change_inv_ds %>%
 # group_by(organization_name) %>%
  #filter(pct_change != Inf & pct_change != -Inf) %>%
#  summarize(mean = mean(pct_change), 
 #           sd = sd(pct_change)) %>%
 # mutate(consistent = ifelse(between(mean + sd, -7, 7) | between(mean - sd, -7, 7), TRUE, FALSE))

## Checking if any value is outside of -7:7, can do with min/max
inv_consist <- pct_change_inv_ds %>%
  group_by(organization_name) %>%
  filter(pct_change != Inf & pct_change != -Inf) %>%
  summarize(min = min(pct_change), 
            max = max(pct_change),
            sd = sd(pct_change)) %>%
  mutate(consistent = ifelse(between(min + sd, -7, 7) & between(max - sd, -7, 7), TRUE, FALSE),
         #Manual inspection showed BQC should be marked as consistent but just didnt quite make it math-wise
         consistent = ifelse(organization_name == "Ballet Quad Cities", TRUE, consistent))

## Adding consistency 
inv_consist_ds <- inv_consist %>% 
  select(organization_name, consistent) %>% 
  right_join(pct_change_inv_ds, by = "organization_name") %>%
  filter(pct_change != Inf & pct_change != -Inf)

```

```{r}
##plottinng
test <- ggplot(inv_consist_ds, aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = consistent)) +
  geom_line(show.legend = TRUE, alpha = 0.3) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance (max 100)",
       subtitle = "By Fiscal Year") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20),
                     limits = c(-100, 100))  +
  geom_point(data = pct_change_inv_med, aes(x = fiscal_year, y = median), color = "black", size = 0.5) + 
  geom_hline(yintercept = 7, alpha = .5, linetype = "dotted") + 
  geom_hline(yintercept = -7, alpha = .5, linetype = "dotted") 

test
ggplotly(test)

inv_box <- inv_consist_ds %>%
  group_by(organization_name) %>%
  filter(pct_change != Inf) %>%
  summarize(sd = sd(pct_change, na.rm = TRUE)) %>%
  right_join(inv_consist_ds, by = "organization_name") %>%
  select(organization_name, EIN, pct_change, sd, consistent) %>%
  mutate(organization_name = reorder(organization_name, -sd, na.rm = TRUE))

## Unlimited 
box_plot <- ggplot(inv_box, aes(x = organization_name, y = pct_change)) + 
  geom_boxplot(aes(color = consistent), show.legend = FALSE) + 
  geom_point(size = 1, alpha = 0.5) + 
  theme_bw() + 
  labs(title = "Range of Endowment Percent Change per Company",
       x = "Dance Company",
       y = "Percentage of Change in Endowment Balance") + 
  theme(axis.text.x = element_blank()) + 
  geom_hline(yintercept = 100, linetype = "dotted", color = "maroon")
ggplotly(box_plot) %>%
  layout(showlegend = FALSE)

##Limited to 100 for visibility 
ggplot(inv_box, aes(x = organization_name, y = pct_change)) + 
  geom_boxplot(aes(color = consistent), show.legend = FALSE) + 
  geom_point(size = 1, alpha = 0.2) + 
  theme_bw() + 
  labs(title = "Range of Endowment % Change (Max of 100) per Company",
       x = "Dance Company",
       y = "Percentage of Change in Endowment Balance") + 
  theme(axis.text.x = element_blank()) + 
  scale_y_continuous(breaks = scales::breaks_pretty(n = 20),
                     limit = c(-100,100))
```

## Examining the Two Classes

### Lists of Companies within Each Class
```{r}
## Names in a table  
# CONSISTENT ones  
inv_consist %>% 
  filter(consistent == TRUE) %>%
  select(-consistent) %>%
  make_table(title = "Companies with Consistent Endowment Changes", col_names = c("Organization Name", "Minimum Change", "Maximum Change", "Standard Deviation")) %>%
  scroll_box(height = "450px")
```
 
There are `r inv_consist %>% filter(consistent == TRUE) %>% nrow()` companies which have 'consistent' endowment percent change. 


```{r}
## Names in a table  
# INCONSISTENT ones  
inv_consist %>% 
  filter(consistent == FALSE) %>%
  select(-consistent) %>%
  make_table(title = "Companies with Inconsistent Endowment Changes", col_names = c("Organization Name", "Minimum Change", "Maximum Change", "Standard Deviation")) %>%
  scroll_box(height = "450px")
```

There are `r inv_consist %>% filter(consistent == FALSE) %>% nrow()` companies which have 'inconsistent' endowment percent change. 

### Size of Companies in Each Class   

```{r}
##Getting employee data
source(here("GET_VARS.R"))
##Specifically reading in employee data
files <- dir( here("ballet_990_released_20230208"),
              full.names = TRUE)
employ_data <- map_df(files, ~
                    get_df(variables =  c("//Return//ReturnData//TotalEmployeeCnt"),
                           filename = .x
                           )) %>%
    mutate(TotalEmployeeCnt = as.numeric(TotalEmployeeCnt)) %>%
  left_join(names, by = "EIN")
```

```{r}
## Most recent years for each company
employ_data$fiscal_year = as.numeric(as.character(employ_data$fiscal_year))
most_recent_yrs <- employ_data %>%
  group_by(organization_name) %>%
  summarize(recent_year = max(fiscal_year, na.rm = TRUE)) %>%
  left_join(employ_data, by = c("organization_name", "recent_year" = "fiscal_year"))

## Limiting to ONLY most recent year 
most_recent_ds <- inv_consist_ds %>%
  left_join(most_recent_yrs, by = "organization_name") %>%
  filter(fiscal_year == recent_year)
```

```{r}
## Consistent  
most_recent_ds %>%
  filter(consistent == TRUE) %>%
  select(organization_name, TotalEmployeeCnt, EndYearBalanceAmt, recent_year) %>%
  make_table("Size of Consistent Companies, by Employees and Endowment", col_names = c("Organization Name", "Total Employee Count", "End Year Balance Amount", "Fiscal Year")) %>%
  scroll_box(height = "450px")
```


```{r}
## Inconsistent  
most_recent_ds %>%
  filter(consistent == FALSE) %>%
  filter(!is.na(TotalEmployeeCnt)) %>%
  select(organization_name, TotalEmployeeCnt, EndYearBalanceAmt, recent_year) %>%
  make_table("Size of Inconsistent Companies, by Employees and Endowment", col_names = c("Organization Name", "Total Employee Count", "End Year Balance Amount", "Fiscal Year")) %>%
  scroll_box(height = "450px")
```

```{r}

## Labels for facet 
consistent_labels <- c(
  `TRUE` = "Consistent",
  `FALSE` = "Inconsistent"
)

##Histograms
most_recent_ds %>%
  ggplot(aes(x = TotalEmployeeCnt)) +
  geom_histogram(binwidth = 60, color = "black", fill = "gray") +
  facet_wrap(~consistent, labeller = as_labeller(consistent_labels)) + 
  theme_bw() + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 90),
        strip.text = element_text(face="bold",size = 10),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
    scale_x_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 10)) +
  labs( title = "Size of Companies by Employee Total")

most_recent_ds %>%
  ggplot(aes(x = EndYearBalanceAmt)) +
  geom_histogram(color = "black", fill = "gray") + 
  facet_wrap(~consistent, labeller = as_labeller(consistent_labels)) + 
  theme_bw() + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 90),
        strip.text = element_text(face="bold",size = 10),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
    scale_x_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 10)) +
  labs( title = "Size of Companies by End Year Balance Amount")

most_recent_ds %>%
  filter(EndYearBalanceAmt < 20000000) %>%
  ggplot(aes(x = EndYearBalanceAmt)) +
  geom_histogram(color = "black", fill = "gray") + 
  facet_wrap(~consistent, labeller = as_labeller(consistent_labels)) + 
  theme_bw() +
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 90),
        strip.text = element_text(face="bold",size = 10),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
    scale_x_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 10)) +
  labs( title = "Size of Companies by End Year Balance Amount, Limit $20,000,000")
```

# Percent Change Within the Pandemic   
## WITH investments 
```{r}  
## Pandemic Years
pct_change_plot <- pct_change_ds %>%
  filter(fiscal_year %in% c("2018", "2019", "2020", "2021", "2022")) %>% 
  ggplot(aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = organization_name)) +
  geom_line(show.legend = FALSE, alpha = 0.5) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance",
       subtitle = "Within Pandemic Years") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20)) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "gray")
ggplotly(pct_change_plot) 
```

```{r}
## Table of available in-pandemic data 
pct_change_ds %>%
  filter(fiscal_year %in% c("2019", "2020", "2021", "2022")) %>%
  select(organization_name, pct_change, fiscal_year) %>%
  arrange(desc(fiscal_year)) %>%
  make_table(title = "Percentage of Change in Endowment Balance within Pandemic Years", col_names = c("Name", "% Change", "Year")) %>%
  scroll_box(height = "450px")
```

```{r}
pct_change_ds %>%
  filter(fiscal_year %in% c("2019", "2020", "2021", "2022")) %>%
  select(organization_name, pct_change, fiscal_year) %>%
  group_by(fiscal_year) %>%
  summarize(total_in_year = n())
```

## WITHOUT investments 
```{r}  
## Pandemic Years
pct_change_plot_inv <- pct_change_inv_ds %>%
  filter(fiscal_year %in% c("2018", "2019", "2020", "2021", "2022")) %>% 
  ggplot(aes(x = fiscal_year, y = pct_change, 
             group = organization_name, color = organization_name)) +
  geom_line(show.legend = FALSE, alpha = 0.5) + 
  theme_bw() + 
  labs(y = "Percent Change",
       x = "Fiscal Year",
       title = "Percentage of Change in Endowment Balance",
       subtitle = "Within Pandemic Years") + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 25),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = scales::pretty_breaks(n = 20)) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "gray")
ggplotly(pct_change_plot_inv) 
```

```{r}
## Table of available in-pandemic data 
pct_change_inv_ds %>%
  filter(fiscal_year %in% c("2019", "2020", "2021", "2022")) %>%
  select(organization_name, pct_change, fiscal_year) %>%
  arrange(desc(fiscal_year)) %>%
  make_table(title = "Percentage of Change in Endowment Balance within Pandemic Years", col_names = c("Name", "% Change", "Year")) %>%
  scroll_box(height = "450px")
```

```{r}
pct_change_inv_ds %>%
  filter(fiscal_year %in% c("2019", "2020", "2021", "2022")) %>%
  select(organization_name, pct_change, fiscal_year) %>%
  group_by(fiscal_year) %>%
  summarize(total_in_year = n())
```

