---
title: "Data Dictionary"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(xml2)
```

## Useful Resources for Reference

* [example form](https://www.irs.gov/pub/irs-pdf/f990sd.pdf)
* [variable definitions w/ reference to which questions they are derived from]( http://www.irsx.info/metadata/forms.html#ReturnHeader990x)

## More Information on Variables


## Reformatting Variable Names to be More Readable for Future Plots and Tables

```{r}

# resources:
# example form: https://www.irs.gov/pub/irs-pdf/f990sd.pdf


# same variables as extracted for the endowment data
variables_no_path <- c("ReturnTs", 
                       "EIN",
                       "DonorRstrOrQuasiEndowmentsInd",
                       "CYBeginningYearBalanceAmt", 
                       "CYContributionsAmt", 
                       "CYInvestmentEarningsOrLossesAmt",
                       "CYOtherExpendituresAmt",
                       "CYEndYearBalanceAmt",
                       
                       "CYM1BeginningYearBalanceAmt", 
                       "CYM1ContributionsAmt", 
                       "CYM1InvestmentEarningsOrLossesAmt",
                       "CYM1OtherExpendituresAmt",
                       "CYM1EndYearBalanceAmt",
                       
                       "CYM2BeginningYearBalanceAmt", 
                       "CYM2ContributionsAmt", 
                       "CYM2InvestmentEarningsOrLossesAmt",
                       "CYM2OtherExpendituresAmt",
                       "CYM2EndYearBalanceAmt",
                       
                       "CYM3BeginningYearBalanceAmt", 
                       "CYM3ContributionsAmt", 
                       "CYM3InvestmentEarningsOrLossesAmt",
                       "CYM3OtherExpendituresAmt",
                       "CYM3EndYearBalanceAmt",
                       
                       "CYM4BeginningYearBalanceAmt", 
                       "CYM4ContributionsAmt", 
                       "CYM4InvestmentEarningsOrLossesAmt",
                       "CYM4OtherExpendituresAmt",
                       "CYM4EndYearBalanceAmt",
                       
                       "BoardDesignatedBalanceEOYPct",
                       "PrmnntEndowmentBalanceEOYPct",
                       "TermEndowmentBalanceEOYPct",
                       "EndowmentsHeldUnrelatedOrgInd",
                       "EndowmentsHeldRelatedOrgInd",
                       
                       "AmendedReturnInd",
                       "ReturnTypeCd"
)


readable_names <- tibble(original_variable = variables_no_path) %>%
  # add clearer interpretation of the time period given
  mutate(time_period = case_when(
    grepl("CYM1", original_variable) ~ "(1 year before current)",
    grepl("CYM3", original_variable) ~ "(2 years before current)",
    grepl("CYM3", original_variable) ~ "(3 years before current)",
    grepl("CY", original_variable) ~ "(current year)"
  )) %>%
  # remove time period since separate column holds that information
  mutate(variable = gsub("CYM1|CYM2|CYM3|CYM4|CY", "", original_variable),
         # split into words where there is capitalization
         var_split = gsub("([[:upper:]])", " \\1", variable),
         # change shortened words to readable format
         readable = gsub("E I N", "EIN", var_split) %>%
           gsub("Rstr", "Restricted",.) %>%
           gsub("Amt", "Amount", .) %>%
           gsub("Ind", "Index", .) %>%
           gsub("Org", "Organization", .) %>%
           gsub("Pct", "Percentage", .) %>%
           gsub("E O Y", "End of Year", .) %>%
           gsub("Prmnnt", "Permanent", .),
         readable = ifelse(!is.na(time_period), 
                           paste(readable, time_period), 
                           readable)) %>% 
  select(original_variable, readable)

saveRDS(readable_names, "./data/readable_names.RDS")      

```

## All Variables Available in Schedule D

To see what variables might be of interest and are well-represented in the dataset, we can look at the variable names and the number of files that have a given variable.

It may be good to get a sense for the variables that are present in the most of our files. The table below ranks the variables extracted by how many files have that variable.

```{r,include = FALSE, eval=FALSE}

############################################
# FUNCTION DEVELOPMENT W/ TEST EXAMPLE
############################################

files <- dir( "./ballet_990_released_20230208",
               full.names = TRUE)

example_file <- xml2::read_xml(files[10])%>% xml_ns_strip()

# look at all variable names in schedule D
IRS990ScheduleD_vars <- example_file %>% 
  xml_find_all("//ReturnData//IRS990ScheduleD")  %>%
  xml_contents() %>% 
  xml_path() 

# look at variable names of children nodes of the schedule D nodes
IRS990ScheduleD_var_children <- IRS990ScheduleD_vars %>% 
  map(~xml_find_all(example_file,
                    xpath = gsub("/", 
                                 "//" ,
                                 .x, 
                                 fixed = TRUE) ) %>%
        xml_children() %>% 
        xml_path()) %>%
  unlist()

# include both first level schedule D nodes and children nodes
vars <- c(IRS990ScheduleD_vars, IRS990ScheduleD_var_children)

```

```{r}
################################################################################
# extract variables in the IRS990ScheduleD node from the given file
################################################################################
# output is a dataframe with the filename and name of each variable extracted

collect_sched_d_vars <- function(input_file) {
  example_file <- xml2::read_xml(input_file)%>% xml_ns_strip()

  # look at all variable names in schedule D
  IRS990ScheduleD_vars <- example_file %>% 
    xml_find_all("//ReturnData//IRS990ScheduleD")  %>%
    xml_contents() %>% 
    xml_path() 
  
  # look at variable names of children nodes of the schedule D nodes
  IRS990ScheduleD_var_children <- IRS990ScheduleD_vars %>% 
    map(~xml_find_all(example_file,
                      xpath = gsub("/", 
                                   "//" ,
                                   .x, 
                                   fixed = TRUE) ) %>%
          xml_children() %>% 
          xml_path()) %>%
    unlist()
  
    # include both first level schedule D nodes and children nodes
    vars <- c(IRS990ScheduleD_vars, IRS990ScheduleD_var_children)
    
    tibble(variables = vars,
           filename = input_file)
}

# file paths in the directory
files <- dir( "./ballet_990_released_20230208",
               full.names = TRUE)

# iterate over all files and extract variables present
# output is a data frame with a file name column and variable name column
all_vars <- map_df(files, collect_sched_d_vars)

```

```{r}

##################################################
# print data frame to show frequency of different 
# variables in the dataset 
##################################################

# version that does not include parent node
# interpretation is more difficult 
all_vars %>% 
   rename(Variable = variables) %>%
   group_by(Variable) %>% 
   summarize(`Number of Files` = n())  %>%
   mutate(Variable = gsub("*.*\\/", "", Variable)) %>%
  arrange(desc(`Number of Files`))


# sort variables by the number of files that have a given variable
# include parent variable for ease of interpretation
all_vars %>% 
   rename(Variable = variables) %>%
   group_by(Variable) %>% 
  # number of files that have each variable
   summarize(files = n()) %>%
  # split variable paths
  #second to last element is parent, last element is variable
   mutate(Variable = strsplit(Variable, split = "/"),
           Parent = map_chr(Variable, ~.x[length(.x)-1]),
           Variable = map_chr(Variable, ~.x[length(.x)])) %>%
  select(Parent, Variable, `Number of Files` = files ) %>%
  arrange(desc(`Number of Files`))


```

