---
title: "Generate Table from XML Files"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{css,echo=FALSE}

#sidebar h2 { background: #3B5561 !important;
}
#main .nav-pills > li.active > a,
#main .nav-pills > li.active > a:hover,
#main .nav-pills > li.active > a:focus {
    background-color: #22983B;
}

#main .nav-pills > li > a:hover {
    background-color: #22983B;
}

h1, h2, h3, h4, h5, h6, legend {
    color: #334B56;
}

#nav-top span.glyphicon {
    color: #22983B;
}



#table-of-contents h2 {
    background-color: #3B5561;
}

#content{
max-width:2300px;
}



```


```{r}

library(tidyverse)
library(xml2)
library(kableExtra)
library(lubridate)
library(forcats)


```

# Generating Table for All Files

```{r,eval=TRUE, class.source = "fold-show" }


##########################################################
# ITERATE THROUGH ALL FILES AND EXTRACT GIVEN VARIABLES
##########################################################


# Reference for Field Names
# https://vtechworks.lib.vt.edu/bitstream/handle/10919/109974/TaxDataFinalReport.pdf?sequence=10&isAllowed=y

# path assumes zip-file is unzipped in smith-capstone-23 directory
files <- dir( "./ballet_990_released_20230208",
              full.names = TRUE)

# variables to extract from each xml file
variables_to_extract <- c("//Return//ReturnHeader//ReturnTs",  
                          "//Return//ReturnHeader//Filer//EIN", 
                          "//Return//ReturnHeader//ReturnTypeCd",
                          "//Return//ReturnHeader//TaxPeriodBeginDt",
                          "//Return//ReturnHeader//TaxPeriodEndDt",
                          "//AmendedReturnInd",
                          "//Return//ReturnHeader//BusinessName")




#' @param variables character vector of variables to extract from the xml files
#' @param filename string containing full path to xml file to be read
#' @param include_paths logical; whether to include paths to extracted variables in data frame

get_df <- function(variables,
                   filename, 
                   include_paths = TRUE) {
  
  # create column names with just the variables (no paths)
  variables_no_path <- gsub("*.*\\/", "", variables )
  
  xml_file <- read_xml(filename)
  xml_file <- xml_ns_strip(xml_file)
  
  # extract each variable; if it isn't present, put NA 
  extracted <- map(variables, ~{
    value <- xml_find_all(
      xml_file, 
      xpath =.x)
    value <- ifelse(length(value) ==0, 
                    NA, 
                    xml_text(value)) })
  
   names(extracted) <- variables_no_path
   
   extracted <- extracted %>%
     as_tibble()
 
 
   # add columns with the paths to each variable within the xml file
   if(include_paths) {
      paths <- map(variables, ~ {
        value <- xml_find_all(
          xml_file, 
          xpath = paste0(.x)) 
         value <- ifelse(length(value) ==0, 
                      NA, 
                      xml_path(value)) })
      
      names(paths) <- paste0(variables_no_path, "_path")
      
      extracted <- extracted %>%
        bind_cols(as_tibble(paths)) 
  
   }
    
   # check how many of the entries are NA
   # if all NA, prefix 'irs:' may be needed
   columns_not_na <- map_dbl(as.data.frame(extracted), ~!is.na(.x)) %>% sum()
   
   # handle case where prefix 'irs' is in front
   if(columns_not_na == 0){
     
       extracted <- map(variables_no_path, ~ {
        value <- xml_find_all(xml_file, 
                              xpath = paste0("//irs:", .x)) 
        value <- ifelse(length(value) ==0, 
                    NA, 
                    xml_text(value)) })
        names(extracted) <- variables_no_path
        
         extracted <- extracted %>%
           as_tibble()
     
        if(include_paths) {
        paths <- map(variables_no_path, ~ {
          value <- xml_find_all(
            xml_file, 
            xpath = paste0("//irs:", .x))
      
          value <- ifelse(length(value) ==0, 
                      NA, 
                      xml_path(value)) })
        
        names(paths) <- paste0(variables_no_path, "_path")  
        
      extracted <-  extracted %>%
        bind_cols(as_tibble(paths)) 
     }
   }
   
 # add name of file to data frame
 extracted %>%
   mutate(filename = gsub("*.*\\/", "", filename))
}



if(!file.exists("./data")) {dir.create("./data")}

all_data <- map_df(files, ~
                     get_df(variables = variables_to_extract,
                     filename = .x)) 

all_data <- all_data %>%
  mutate(ReturnDate = as.Date(ReturnTs,
                              format = "%Y-%m-%d")) %>%
  mutate(across(c(TaxPeriodBeginDt, 
                  TaxPeriodEndDt),
                as.Date, 
                format = "%Y-%m-%d")) %>%
  mutate(fiscal_year = year(TaxPeriodEndDt),
         fiscal_year = factor(fiscal_year),
         BusinessName = toupper(BusinessName))

# check all files are present in data frame
# nrow(all_data) == length(files)

saveRDS(all_data, "./data/data_990.RDS")

```


```{r troubleshooting, include=FALSE,eval=FALSE}

# some files have irs prefix; this needs special handling

filename <-  "./ballet_990_released_20230208/201940149349301304_public.xml"
test <- get_df(filename = filename, variables=variables_to_extract)

```

```{r function to generate table}

make_table <- function(..., title = "") {
  title <- paste0("<center><span style = 'font-size:150%;color:black'><b>",
                  title,
                  "</span></b><center>")
   as_tibble(...) %>%
    kbl(caption = title) %>%
    kable_material() %>%
    row_spec(row=0, background = "#43494C" , color = "white", bold = TRUE)
}


```

```{r load data}

# load table of data from parsed xml files

all_data <- readRDS("./data/data_990.RDS")




```

# 1. Filings per Year

## Filings by Fiscal Year

```{r}


############################################################
# TABLE: Number of Filings by Year of Return Date
############################################################
all_data %>%
  group_by(fiscal_year) %>%
  summarize(`Number of Filings`= n()) %>%
  rename(`Year of Return Date` = fiscal_year) %>%
  make_table(title = "Number of Filings by Fiscal Year")


```

```{r, fig.width = 10, fig.height = 6}

############################################################
# PLOT: Number of Filings by Year of Return Date
############################################################
all_data %>%
  group_by(fiscal_year) %>%
  summarize(`Number of Filings`= n()) %>%
  rename(`Fiscal Year` = fiscal_year) %>%
  ggplot(aes(x =`Fiscal Year`, y = `Number of Filings`)) +
  geom_bar(stat = "identity", fill = "#3D6071") +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        axis.text.x = element_text(size = 15),
        axis.text.y=element_text(size = 12)) +
  labs(title = "Number of Filings by Fiscal Year") +
  scale_y_continuous(n.breaks = 6)


```

# 2. Companies per Year

```{r, fig.width = 10, fig.height = 6}


####################################################
# TABLE OF (DISTINCT) EINS PER YEAR
####################################################
all_data %>%
  group_by(fiscal_year) %>%
  summarize(`Number of Companies` = n_distinct(EIN)) %>%
  rename(`Fiscal Year` = fiscal_year) %>%
  make_table(title = "Number of Companies by Fiscal Year")


####################################################
# PLOT OF (DISTINCT) EINS PER YEAR
####################################################
all_data %>%
  group_by(fiscal_year) %>%
  summarize(`Number of Distinct EINs` = n_distinct(EIN)) %>%
  rename(`Fiscal Year` = fiscal_year) %>%
  ggplot(aes(x = `Fiscal Year`, y = `Number of Distinct EINs`)) +
  geom_bar(stat = "identity", fill = "#3D6071") +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        axis.text.x = element_text(size = 15)) +
  labs(title = "Number of Companies by Fiscal Year") 
     


```

# 3. Companies with More than One Filing in a Year

We can see the EINs and names of the companies with more than one filing in some fiscal year in the table below.

```{r}
############################################################
# TABLE WITH COMPANIES WITH MORE THAN ONE FILING, BY YEAR
############################################################
all_data %>%
  group_by(fiscal_year, EIN) %>%
  # number of filings for each company for each year
  summarize(n_filings =n(),.groups="drop") %>%
  filter(n_filings > 1) %>%
  rename(`Fiscal Year` = fiscal_year,
         `Number of Filings` = n_filings) %>%
  arrange( `Fiscal Year`, desc(`Number of Filings`)) %>%
  make_table(
    title = "Companies with More than One Filing<br>by Fiscal Year") %>%
  scroll_box(height = "450px")

```

<br><br><br>


We can also look at the trends in the number of companies with 1 filing or 2 filings by  fiscal year in the plot below.


```{r, fig.width = 12, fig.height = 6}


all_data %>%
  group_by(fiscal_year, EIN) %>%
  # number of filings for each company for each year
  summarize(n_filings =n(),.groups="drop") %>% 
  group_by(n_filings, fiscal_year) %>%
  # number of companies with n_filings for each year
  summarize(n_companies = n(), .groups="drop") %>%
  mutate(n_filings = factor(n_filings)) %>%
  ggplot(aes(x = n_filings, 
             y =n_companies,
             fill = n_filings)) +
  geom_bar(stat = "identity",
           position = "dodge",
           show.legend=FALSE) +
  facet_wrap(~fiscal_year, nrow =1) +
  scale_fill_manual(
    values = c("#72e5ef",
               "#1e438d", 
               "#58df8c")) +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 15),
        strip.text = element_text(face="bold",size = 18)) +
  labs(y = "Number of Companies",
       x = "Number of Filings",
       title = "Companies by Number of Filings per Year",
       subtitle = "By Fiscal Year")

```


# 4. Gaps in Filings 

## Gaps in Filings for All Companies {.tabset}

### By EIN 

```{r, fig.height = 16,  fig.width = 10}

all_data %>% 
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  mutate(sum_coverage = sum(time))%>%
  ggplot(aes(xmin = TaxPeriodBeginDt, 
             xmax = TaxPeriodEndDt, 
             y = fct_reorder(EIN, sum_coverage))) +
  geom_linerange(size = 1, alpha = .8) +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11,angle=20, vjust = .7),
        axis.text.y = element_text(size = 4),
        strip.text = element_text(face="bold",size = 18)) +
  scale_x_date(date_breaks = "12 months",
               date_labels = "%b %Y") +
  labs(y = "EIN",
       x = "Date",
       title = "Gaps in Coverage by EIN")
  # scale_x_date(date_breaks = "9 months",
  #              date_labels = "%b %Y") +

```

### By Company Name


```{r, fig.height = 16,  fig.width = 10}


ein_to_name <- all_data %>%
  select(EIN, BusinessName) %>%
  group_by(EIN) %>%
  mutate(n_names = n_distinct(BusinessName)) %>%
  slice_head(n=1) %>%
  select(EIN, BusinessName)


all_data %>% 
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  mutate(sum_coverage = sum(time))%>%
  left_join(ein_to_name) %>%
  ggplot(aes(xmin = TaxPeriodBeginDt, 
             xmax = TaxPeriodEndDt, 
             y = fct_reorder(BusinessName, sum_coverage))) +
  geom_linerange(size = 1, alpha = .8) +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11,angle=20, vjust = .7),
        axis.text.y = element_text(size = 4),
        strip.text = element_text(face="bold",size = 18)) +
  scale_x_date(date_breaks = "12 months",
               date_labels = "%b %Y") +
  labs(y = "Business Name",
       x = "Date",
       title = "Gaps in Coverage by Business Name") 
  # scale_x_date(date_breaks = "9 months",
  #              date_labels = "%b %Y") +

```


## Companies with at Least One Gap Year

To check for gap years, we can take the difference between `TaxPeriodBeginDt` and the lagged `TaxPeriodEndDt` for each EIN.

```{r}

###################################################
# TABLE WITH COMPANIES WITH AT LEAST ONE GAP YEAR
###################################################
all_data %>% 
  # do not include ammended returns for this computation
  filter(is.na(AmendedReturnInd) & ReturnTypeCd != "990T") %>%
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  group_by(EIN) %>%
  arrange(TaxPeriodBeginDt) %>%
  mutate(before =  lag(TaxPeriodEndDt, n =1),
        gap = interval(before, TaxPeriodBeginDt),
        gap = as.numeric(gap, "days")) %>%
  filter(gap >=365)  %>% 
  select(EIN,
         `Previous End of Fiscal Year` = before,
         `Beginning of Fiscal Year` = TaxPeriodBeginDt,
         `Gap Between Filings (Days)` = gap) %>%
  make_table(title = "Companies with a Gap of One Year or More Between Filings") %>%
  scroll_box(height = "450px")

# vector of EINs corresponding to companies with at least one gap year
ein_with_gap <- all_data %>% 
  # do not include ammended returns for this computation
  filter(is.na(AmendedReturnInd) & ReturnTypeCd != "990T") %>%
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  group_by(EIN) %>%
  arrange(TaxPeriodBeginDt) %>%
  mutate(before =  lag(TaxPeriodEndDt, n =1),
        gap = interval(before, TaxPeriodBeginDt),
        gap = as.numeric(gap, "days")) %>%
  filter(gap >=365) %>%
  pull(EIN)

```


### Plotting Gaps in Coverage for Companies with at Least One Gap Year {.tabset}
 
#### By EIN

```{r, fig.width = 11}

############################################################
# PLOT COVERAGE FOR COMPANIES WITH 1+ GAP YEAR (BY EIN)
############################################################
all_data %>% 
  filter(EIN %in% ein_with_gap) %>%
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  mutate(sum_coverage = sum(time))%>%
  ggplot(aes(xmin = TaxPeriodBeginDt, 
             xmax = TaxPeriodEndDt, 
             y = fct_reorder(EIN, sum_coverage))) +
  geom_linerange(size = 1, alpha = .8) +
  theme_bw() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11,angle=20, vjust = .7),
        axis.text.y = element_text(size = 8),
        strip.text = element_text(face="bold",size = 18)) +
  scale_x_date(date_breaks = "12 months",
               date_labels = "%b %Y") +
  labs(y = "EIN",
       x = "Date",
       title = "Gaps in Coverage for Companies\nwith at Least One Gap Year")

```

#### By Business Name

```{r, fig.width = 11}

############################################################
# PLOT COVERAGE FOR COMPANIES WITH 1+ GAP YEAR (BY NAME)
############################################################
all_data %>% 
  filter(EIN %in% ein_with_gap) %>%
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  mutate(sum_coverage = sum(time))%>%
  left_join(ein_to_name) %>%
  ggplot(aes(xmin = TaxPeriodBeginDt, 
             xmax = TaxPeriodEndDt, 
             y = fct_reorder(BusinessName, sum_coverage))) +
  geom_linerange(size = 1, alpha = .8) +
  theme_bw() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11,angle=20, vjust = .7),
        axis.text.y = element_text(size = 8),
        strip.text = element_text(face="bold",size = 18)) +
  scale_x_date(date_breaks = "12 months",
               date_labels = "%b %Y") +
  labs(y = "EIN",
       x = "Date",
       title = "Gaps in Coverage for Companies\nwith at Least One Gap Year")

```


## Distribution of Gaps Between Filings

```{r}
#####################################
# HISTOGRAM OF GAPS BETWEEN FILINGS
#####################################
all_data %>% 
  # do not include ammended returns for this computation
  filter(is.na(AmendedReturnInd) & ReturnTypeCd != "990T") %>%
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  group_by(EIN) %>%
  arrange(TaxPeriodBeginDt) %>%
  mutate(before =  lag(TaxPeriodEndDt, n =1),
        gap = interval(before, TaxPeriodBeginDt),
        gap = as.numeric(gap, "days")) %>%
  ggplot(aes(x =gap )) +
  geom_histogram( fill = "#3D6071") +
  theme_bw() +
  labs(x = "Gap Between Filings (Days)",
       y = "Number of Companies",
       title = "Distribution of the Gap Between Filings") +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11, vjust = .7),
        axis.text.y = element_text(size = 12),
        strip.text = element_text(face="bold",size = 18)) +
  scale_x_continuous(n.breaks = 10)





```

