---
title: "Retrieve Endowment Funds"
author: "Rose Evard"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
---

```{r}

library(tidyverse)
library(xml2)
library(kableExtra)
library(lubridate)
library(forcats)


```


## Extracting Endowments  
```{r extracting endowment information}

##########################################################
# ITERATE THROUGH ALL FILES AND ENDOWMENT  VARIABLES
##########################################################

#' @param filename string containing full path to xml file to be read

##Based off of Quinn's strategy to extract metadata
get_endowment <- function(filename) {
  
  # Retrieving the same endowment information for all 
  variables <- c("//Return//ReturnHeader//ReturnTs", 
                 "//Return//ReturnHeader//Filer//EIN", 
                 "//Return//ReturnData//IRS990//DonorRstrOrQuasiEndowmentsInd",
                 "//Return//ReturnData//IRS990ScheduleD//CYEndwmtFundGrp//BeginningYearBalanceAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYEndwmtFundGrp//ContributionsAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYEndwmtFundGrp//InvestmentEarningsOrLossesAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYEndwmtFundGrp//OtherExpendituresAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYEndwmtFundGrp//EndYearBalanceAmt",
                 
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus1YrEndwmtFundGrp//BeginningYearBalanceAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus1YrEndwmtFundGrp//ContributionsAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus1YrEndwmtFundGrp//InvestmentEarningsOrLossesAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus1YrEndwmtFundGrp//OtherExpendituresAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus1YrEndwmtFundGrp//EndYearBalanceAmt",
                
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus2YrEndwmtFundGrp//BeginningYearBalanceAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus2YrEndwmtFundGrp//ContributionsAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus2YrEndwmtFundGrp//InvestmentEarningsOrLossesAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus2YrEndwmtFundGrp//OtherExpendituresAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus2YrEndwmtFundGrp//EndYearBalanceAmt",
                 
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus3YrEndwmtFundGrp//BeginningYearBalanceAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus3YrEndwmtFundGrp//ContributionsAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus3YrEndwmtFundGrp//InvestmentEarningsOrLossesAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus3YrEndwmtFundGrp//OtherExpendituresAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus3YrEndwmtFundGrp//EndYearBalanceAmt",
                 
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus4YrEndwmtFundGrp//BeginningYearBalanceAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus4YrEndwmtFundGrp//ContributionsAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus4YrEndwmtFundGrp//InvestmentEarningsOrLossesAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus4YrEndwmtFundGrp//OtherExpendituresAmt",
                 "//Return//ReturnData//IRS990ScheduleD//CYMinus4YrEndwmtFundGrp//EndYearBalanceAmt",
                 
                 "//Return//ReturnData//IRS990ScheduleD//BoardDesignatedBalanceEOYPct",
                 "//Return//ReturnData//IRS990ScheduleD//PrmnntEndowmentBalanceEOYPct",
                 "//Return//ReturnData//IRS990ScheduleD//TermEndowmentBalanceEOYPct",
                 "//Return//ReturnData//IRS990ScheduleD//EndowmentsHeldUnrelatedOrgInd",
                 "//Return//ReturnData//IRS990ScheduleD//EndowmentsHeldRelatedOrgInd",
                 
                 "//AmendedReturnInd",
                 "//Return//ReturnHeader//ReturnTypeCd"
                 )
  
  # Column name; order matters, needs to align with retrieval order
  variables_no_path <- c("ReturnTs", 
                         "EIN",
                         "DonorRstrOrQuasiEndowmentsInd",
                        "CYBeginningYearBalanceAmt", 
                        "CYContributionsAmt", 
                        "CYInvestmentEarningsOrLossesAmt",
                        "CYOtherExpendituresAmt",
                        "CYEndYearBalanceAmt",
                        
                        "CYM1BeginningYearBalanceAmt", 
                        "CYM1ContributionsAmt", 
                        "CYM1InvestmentEarningsOrLossesAmt",
                        "CYM1OtherExpendituresAmt",
                        "CYM1EndYearBalanceAmt",
                        
                        "CYM2BeginningYearBalanceAmt", 
                        "CYM2ContributionsAmt", 
                        "CYM2InvestmentEarningsOrLossesAmt",
                        "CYM2OtherExpendituresAmt",
                        "CYM2EndYearBalanceAmt",
                        
                        "CYM3BeginningYearBalanceAmt", 
                        "CYM3ContributionsAmt", 
                        "CYM3InvestmentEarningsOrLossesAmt",
                        "CYM3OtherExpendituresAmt",
                        "CYM3EndYearBalanceAmt",
                        
                        "CYM4BeginningYearBalanceAmt", 
                        "CYM4ContributionsAmt", 
                        "CYM4InvestmentEarningsOrLossesAmt",
                        "CYM4OtherExpendituresAmt",
                        "CYM4EndYearBalanceAmt",
                        
                        "BoardDesignatedBalanceEOYPct",
                        "PrmnntEndowmentBalanceEOYPct",
                        "TermEndowmentBalanceEOYPct",
                        "EndowmentsHeldUnrelatedOrgInd",
                        "EndowmentsHeldRelatedOrgInd",
                        
                        "AmendedReturnInd",
                        "ReturnTypeCd"
                        )
  
  xml_file <- read_xml(filename)
  xml_file <- xml_ns_strip(xml_file)
  
  # extract each variable; if it isn't present, put NA 
  extracted <- map(variables, ~{
    value <- xml_find_all(
      xml_file, 
      xpath =.x)
    value <- ifelse(length(value) ==0, 
                    NA, 
                    xml_text(value)) })
  
   names(extracted) <- variables_no_path
   
   extracted <- extracted %>%
     as_tibble()
}
```

```{r}
test_path <- "nycb_fy2020.xml"
test <- get_endowment(test_path)
```

> An important question: Does ZERO in a boolean situation mean false or mean NA?  There are cases where 1s and 0s are coded for T/F.

```{r}
##Applying get_endowment to entire output 
files <- dir( "./ballet_990_released_20230208",
              full.names = TRUE)

endowment_data <- map_df(files, ~
                     get_endowment(.x)) 

## Adjusting data type
endowment_data <- endowment_data %>%
  mutate(ReturnDate = as.Date(ReturnTs,
                              format = "%Y-%m-%d")) %>%
  mutate(across(CYBeginningYearBalanceAmt:TermEndowmentBalanceEOYPct,
                as.numeric)) %>%
  mutate(across(c(EndowmentsHeldRelatedOrgInd, EndowmentsHeldUnrelatedOrgInd, DonorRstrOrQuasiEndowmentsInd),
                ~ifelse(.x == "true" | .x == "1", TRUE, FALSE))) %>%
  filter(is.na(AmendedReturnInd)) %>% #Filtering out data
  filter(ReturnTypeCd == "990" | ReturnTypeCd == "990EZ") %>%
  select(-c(ReturnTypeCd,AmendedReturnInd)) #Removing columns needed for filtering
```

```{r saving data}
saveRDS(endowment_data, "./data/endowment_data_990.RDS")
```

```{r}
##Graphing difference between beginning and end periods
endowment_data %>%
  select(ReturnDate, EIN, CYBeginningYearBalanceAmt, CYEndYearBalanceAmt) %>%
  pivot_longer(cols = c(CYBeginningYearBalanceAmt, CYEndYearBalanceAmt), names_to = "Period", values_to = "Balance") %>%
  mutate(Period = as.factor(Period)) %>%
  mutate(Year = substr(ReturnDate, 1, 4)) %>%
  ggplot(aes(x = Period, y = Balance, group = EIN)) +
  facet_wrap(~Year) +
  geom_point(alpha = 0.5) + 
  geom_line(alpha = 0.5) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 35, hjust =1),
        legend.position = "none") +
  labs(title = "Comparing Fiscal Year Beginning and End Endowment Totals") +
  scale_y_continuous(labels = scales::label_dollar())
```


## Determining Which Needed to File Schedule D for Endowment

Not all those that fill out schedule D do so for endowment, so I focused on Question 10 "Did the organization, directly or through a related organization, hold assets in temporarily restricted endowments, permanent endowments, or quasi endowments?  If "yes," complete Schedule D, Part V".  

```{r}
##Identifying those that DID fill out schedule D for Endowment, as they supplied info in any numeric column
##Assuming the sum of from cols 4-31 (schedule D finances) is not zero
fill_d <- endowment_data %>%
 mutate(FilledSchD = ifelse(rowSums(across(CYBeginningYearBalanceAmt:TermEndowmentBalanceEOYPct), na.rm = TRUE) != 0, TRUE, FALSE)) %>%
  select(ReturnDate, EIN, DonorRstrOrQuasiEndowmentsInd, FilledSchD)

##Counting those that needed to fill schedule D
endowment_data %>%
  group_by(DonorRstrOrQuasiEndowmentsInd) %>%
  summarize(n = n()) %>%
  kbl(caption = "Total EINs checked YES on Q10") %>%
  kable_material()

##Seeing those that DID and or DID NOT need to fill out Schedule D, yet Did or Did Not
fill_d %>%
  group_by(DonorRstrOrQuasiEndowmentsInd, FilledSchD) %>%
  summarize(n = n()) %>%
  kbl(caption = "990 Q10 answer and If They Filled out Schedule D") %>%
  kable_material()
```

> Editor's note: Prior to filtering the data for 990T and amended files, there were ones that marked Q10 as true, approximately 20 of them.  
